Article Title

### 2013-07-24 sections/Abstract.tex

Linear correlations between the spiking activity of pairs of neurons are among the most familiar and useful descriptive statistics of neural activity.
Multineuronal recordings provide far richer information than the equivalent number of neuron pairs considered in isolation.
For example, the full covariance matrix reveals the partial correlations between pairs of neurons or the common correlated activity involving the entire population.
Usual estimates of covariance matrix parameters require large samples for convergence.
Convergence can be improved by regularization, i.e.~by imposing some kind of structure.  
The optimal choice of the regularization approach is largely an empirical question as it depends profoundly on how well the imposed structure can still capture the dominant underlying physiogical interactions.
For example, presumed diffuse correlated input onto many neurons motivate low-rank parameterizations such as in factor analysis.  
Conversely, presumed strong pairwise linear associations between neurons motivate regularization by sparsifying the inverse covariance matrix.
Finally, `sparse+low-rank' inverse covariance parameterizations allow combining diffuse correlations and strong pairwise associations.

To infer the optimal structure of neural correlations in a specitifc neural microcircuit, we compared the performance of several covariance estimators on the activity of 100--300 neurons in mouse visual cortex: sample covariance, Ledoit-Wolf covariance shrinkage, factor analysis, sparse inverse covariance, and sparse+low-rank inverse covariance.
The instantaneous firing rates, binned at 200 ms, were inferred from the somatic calcium signals acquired with fast 3D random-access two-photon microscopy.   
Each covariance estimator was optimized and evaluated by cross-validation.
Unsurprisingly, covariance shrinkage reliably outperformed the sample covariance estimate.
In turn, factor analysis-based estimates significantly outperformed covariance shrinkage.
Yet sparse inverse covariance with or without an additional low-rank component significantly outperformed both factor analysis and shrinkage  estimators.
The superior performance of the sparse inverse covariance estimator suggests the relative importance of detailed network interactions over common diffuse input in this circuit, although these results may vary in other circuits or in other brain states.

>>>>

Linear correlations between the spiking activity of pairs of neurons are among the most familiar and useful descriptive statistics of neural activity. Multineuronal recordings provide far richer information than the equivalent number of neuron pairs considered in isolation. For example, the full covariance matrix reveals the partial correlations between pairs of neurons, as well as correlated activity across the entire population. Covariance matrix estimates require large samples for convergence. Convergence  can be improved by regularization, \emph{i.e.}~by imposing some kind of structure.  The optimal regularization must be determined empirically since its performance depends on how the imposed structure captures the underlying interactions. For example, we can use low-rank parameterizations of the covariance matrix to account for correlated input onto many neurons. Conversely, if correlations are a result of a small number of  pairwise linear interactions between the observed neurons, we can sparsify the inverse covariance matrix. We can also use a `sparse+low rank’ inverse covariance representation to account for common fluctuations and pairwise interactions in the recorded population.

To select the optimal structure of neural correlations in a local neural circuit, we compared the performance of several covariance estimators on the activity of 100--300 neurons in mouse visual cortex: sample covariance, Ledoit-Wolf covariance shrinkage, factor analysis, sparse inverse covariance, and sparse+low-rank inverse covariance. We inferred instantaneous firing rates in 200 ms bins from the somatic calcium signals acquired with fast 3D random-access two-photon microscopy.  Each covariance estimator was optimized and evaluated by cross-validation. As expected, covariance shrinkage reliably outperformed the sample covariance estimate. In turn, factor analysis-based estimates significantly outperformed covariance shrinkage. Yet sparse inverse covariance with or without an additional low-rank component significantly outperformed both factor analysis and shrinkage estimators. The superior performance of the sparse inverse covariance estimator suggests the relative importance of detailed network interactions over common diffuse input in the circuit we studied.

### 2013-07-24 sections/Introduction.tex

\hl{\tiny why correlations:}
Linear correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \citep{Cohen:2011}.   
For example, \emph{noise correlations}, i.e.~the correlations of stimulus response variability between pairs of neurons in sensory areas, have been of particular interest thanks, in part, to their profound theoretical implications for stimulus coding \citep{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011}.
Lending further credence to their role as indicators of detailed and specific functional organization have been a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons, their synaptic connectivity and stimulus tuning similarity, cortical layer specificity, cell-type specificity, progressive changes in development and in learning, changes due to sensory stimulation and global brain states, and others \citep{Kohn:2005,Smith:2008,Kohn:2009,Goard:2009,Golshani:2009,Renart:2010,Ecker:2010,Smith:2013,Denman:2013}.

\hl{\tiny ambiguity of interpretation:}
Neural correlations do not come with ready or unambiguous mechanistic interpretations.   
Theoretical and simulation studies show that neural correlations at various temporal scales may arise from any combination of underlying mechanisms such as direct synaptic interactions,  common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \citep{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}

\hl{\tiny opportunities with multineuronal data:}
Early studies of neural correlations were based on measurements from isolated pairs of neurons and their impact on coding was extrapolated to entire populations \citep{Shadlen:1998,Zohary:1994}.  Modern massively multineuronal recordings allow estimations of entire correlation matrices from large neuronal populations, which provide more information than the equivalent number of pairwise correlations assessed in isolation.
After all, the correlation matrix is greater than the sum of its parts: it can be transformed into other representations  that accentuate different aspects of its structure and may suggest different mechanistic interpretations.  
For example, the eigenvalue decomposition, also known as the principal component analysis, of the covariance matrix expresses shared correlated activity components across the entire population;  common fluctuations of population activity of the entire circuit may be accurately represented by just a few principal components but will affect all correlation coefficients. 
In contrast, the off-diagonal elements of the inverse of the correlation matrix constitute scaled partial correlations between neuron pairs, which reflect their specific linear dependencies, after accounting for all other recorded cells;
a strong interaction between a pair of neurons may be expressly represented by a single partial correlation but its effects propagate to multiple correlations and principal components.

\hl{\tiny motivate regularization} Estimation of parameters of the covariance matrix from data is inherently challenging, increasingly for very high-dimensional data. As the amount of data grows only linearly with population size, the number of free parameters in a complete representation of the covariance matrix increases quadratically, multiplying opportunities for spurious patterns to emerge among the correlations. 
Although the estimation error of each covariance coefficient, in isolation, is unaffected by the dimensionality of the data, the estimation error of other parameters of the covariance matrix increases with population size.  
For example, the large eigenvalues of in the sample covariance matrix are biased upward while small eigenvalues are biased downward, with both variance and bias increasingly large with increasing data dimensionality \citep{Hayes:1981}.  Similarly, the coefficients of the inverse covariance matrix require  larger sample sizes to be estimated accurately in high-dimensional data.

\hl{\tiny forms of regularization} The convergence of the estimate of multivariate parameters to their true values can be improved through \emph{regularization},  Regularization is the deliberate biasing or \emph{shrinkage} of the empirical estimate toward a less variable \emph{target estimate} \citep{Schafer:2005} or the deliberate restriction of the estimate to a lower-dimensional space \citep{Dempster:1972,Fan:2008,Friedman:2008,Rothman:2008}. 
\hl{\tiny The two definitions overlap and are often equivalent, but not always, according to James, but I would like to see an example of shrinkage that is not, effectively, a low dimensional parameterization of the original problem.} 
One intuitive explanation of regularization is as the addition of some  prior knowledge about the likely forms of the covariance matrix. Curiously, though, regularization need not reflect any accurate knowledge about the nature of covariance matrices in general or in the particular domain. Some improvement arises due to \emph{Stein's phenomenon}, i.e. the favorable tradeoff between bias and variability \emph{anytime} a moderate bias toward a less variable target is introduced.  
However, the incorporation into the regularization scheme of real knowledge about the probable forms of covariance matrices for a specific situation will likely confer additional advantage and improve the convergence more substantially than an \emph{ad hoc} regularizer.

%Estimator improvements due to regularization depend profoundly on the parameterization to which it is applied.  Sparsification of the eigenspectrum produces a low-rank estimate of the covariance matrix also known as truncated PCA \citep{Rothman:2008} or, allowing for additional independent variances, as factor analysis (FA) \citep{Fan:2008}.  Sparsification of the inverse covariance is known as covariance selection \citep{Dempster:1972,Friedman:2008} and is related to finding a graphical model in which a large fraction of neuron pairs are conditionally independent of each other.  Representations that are capable of concentrating the significant real effects in a small number of parameters will reap greatest benefits from sparsification over ones that distribute the real effects widelyacross many parameters.  Random noise tends to be broadly distributed across all parameters in any parameterization.

Two questions must be answered in the choice of the regularization scheme: what kind of regularization and how much regularization. The answer to the first question is domain-specific and is often motivated by theories of the origins of correlations in the specific subject domain.

\hl{\tiny approach and summary of findings} 
In this study, we compared the performance of several covariance estimation schemes for spatially compact groups of 150--300 neurons in layers 2-4 in mouse primary visual cortex during visual stimulation.   The performance of the estimators was evaluted by computing the mean squared error between the optimized covariance estimate fitted to training data and the non-regularized covariance matrix from a separate test data set.  Low-rank covariance estimates performed significantly better than shrinkage estimators, but estimators with sparse partial correlations were more efficient still. Typically, between 3 and 16\% of neuronal paris were connected by non-zero partial interactions.  Mixed sparse estimators with sparse partial correlations and low-rank components performed comparably.

As noted above, the optimal covariance estimator is domain-dependent. 
These specific findings may not spread to all neural circuits.
Improved accuracy estimation accuracy is not the main aim or the only benefit.  
The selection of the optimal regularization scheme in itself serves as an effective descriptor of the dominant correlation structures. The finding that a identifiable subsets of partial correlations were important to describe the partial correlation makes such pairs as prime candidates as direct functional interactions.

>>>>

\hl{\tiny why correlations:} Linear correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \citep{Cohen:2011}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been of particular interest.  Noise correlations can have profound theoretical implications for stimulus coding \citep{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011}, and have been interpreted to indicate detailed and specific functional organization. Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distance separating the neurons, their synaptic connectivity and stimulus tuning similarity, cortical layer specificity, cell-type specificity, progressive changes in development and in learning, changes due to sensory stimulation and global brain states, and others \citep{Kohn:2005,Smith:2008,Kohn:2009,Goard:2009,Golshani:2009,Renart:2010,Ecker:2010,Smith:2013,Denman:2013}.

\hl{\tiny ambiguity of interpretation:} Neural correlations do not come with ready or unambiguous mechanistic interpretations.  Theoretical and simulation studies show that neural correlations at various temporal scales may arise from any combination of underlying mechanisms.  These include direct synaptic interactions, common or correlated inputs, shared sensory noise, chains of multiple synaptic connections, oscillations, top-down modulation, and background network activity \citep{Perkel:1967b,Shadlen:1998,Salinas:2001,Ostojic:2009,Rosenbaum:2011}

\hl{\tiny opportunities with multineuronal data:} Early studies of neural correlations were based on measurements from isolated pairs of neurons and their impact on coding was extrapolated to entire populations \citep{Shadlen:1998,Zohary:1994}.  Multineuronal recordings allow estimation of covariance matrices of larger populations.  Such estimates provide more information than the equivalent number of pairwise correlations assessed in isolation. Indeed, the correlation matrix is greater than the sum of its parts: it can be transformed into other representations that accentuate different aspects of the correlation structure and may suggest different mechanistic interpretations. For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients. In contrast, the off-diagonal elements of the inverse of the correlation matrix constitute scaled partial correlations between neuron pairs, which reflect their specific linear dependencies, after accounting for the activity of all the other recorded cells; a strong interaction between a pair of neurons may be expressed by a single partial correlation but its effects may propagate to multiple correlations and eigenvalues.

\hl{\tiny motivate regularization} Estimation of parameters of the covariance matrix from data is inherently challenging, increasingly so for very high-dimensional data. As the amount of data grows only linearly with population size, the number of free parameters in a complete representation of the covariance matrix increases quadratically, multiplying opportunities for spurious patterns to emerge among the correlations.  Although the estimation error of each covariance coefficient, in isolation, is unaffected by the dimensionality of the data, the estimation error of other parameters of the covariance matrix increases with population size.  
For example, the large eigenvalues of in the sample covariance matrix are biased upward while small eigenvalues are biased downward, with both variance and bias increasingly large with increasing data dimensionality \citep{Hayes:1981}.  Similarly, the coefficients of the inverse covariance matrix require larger sample sizes to be estimated accurately in high-dimensional data.

\hl{\tiny forms of regularization} The convergence of the estimate of multivariate parameters to their true values can be improved through \emph{regularization},  Regularization is the deliberate biasing or \emph{shrinkage} of the empirical estimate toward a less variable \emph{target estimate} \citep{Schafer:2005} or the deliberate restriction of the estimate to a lower-dimensional space \citep{Dempster:1972,Fan:2008,Friedman:2008,Rothman:2008}.  A common intuitive explanation for the effect of regularization is that regularization adds some prior knowledge about the likely forms of the covariance matrix. Curiously, though, regularization need not reflect any plausible knowledge of the nature of covariance matrices in general or in the particular domain. Convergence can improve due to a purely statistical effect, sometimes known as \emph{Stein's phenomenon}, \emph{i.e.}~the favorable tradeoff between bias and variability, which arises \emph{anytime} a moderate bias toward a less variable target is mixed into an unbiased estimate.  Therefore, an estimator that is shown to converge faster on specific types of data does not necessarily contain within itself some wisdom of the statistical regularities in the data.  However, an estimator that does happen to accommodate the statistical structure of the data in a specific domain, it will gain additional advantage over an \emph{ad hoc} regularizer for that domain. Thus the optimal selection of the estimator for a specific problem with unknown statistical structure becomes an empirical question. And, if a clear winner emerges from a number of strong candidates, its regularization scheme can server as an effective statistical description of the covariance structure.

\hl{\tiny approach and summary of findings} In this study, we compared the performance of several covariance estimators for spatially compact groups of 150--300 neurons in layers 2--4 in mouse primary visual cortex during visual stimulation.   The performance of the estimators was evaluated by computing the mean squared error between the optimized covariance estimate fitted to training data and the non-regularized covariance matrix from a separate test data set.  Low-rank covariance estimates performed significantly better than shrinkage estimators, but estimators with sparse partial correlations were more efficient still. Typically, between 3 and 16\% of neuronal pairs were connected by non-zero partial interactions.  Mixed sparse estimators with sparse partial correlations and low-rank components performed comparably.

\hl{\tiny summary of conlusions} As noted above, the optimal covariance estimator is domain-dependent. These specific findings may not spread to all neural circuits.
Improved accuracy estimation accuracy is not the main aim or the only benefit.  
The selection of the optimal regularization scheme in itself serves as an effective descriptor of the dominant correlation structures. The finding that identifiable subsets of partial correlations were important to describe the partial correlation makes such pairs as prime candidates as direct functional interactions.
